---
title: "HW9_2"
output: html_document
---
# Assingment

  Answer the following questions:   
  
Q1)  Can some form of bagging help in your research project?    
    
    
    Yes, My research is about Actibvity Recognition (AR) using different classifiers, and of course bagging could be a very good classifier for the AR problem. 
    

Q2) If it can apply bagging to your research project? Does it help?     

Here, I will apply Bagging method to classify the activity recognition dataset. 


1) Data Prepration

```{r results="hide"}

dt1<-read.table("/Users/mohsennabian/Desktop/MHEALTHDATASET/mHealth_subject1.log")
summary(dt1)
dim(dt1)

dt2<-read.table("/Users/mohsennabian/Desktop/MHEALTHDATASET/mHealth_subject2.log")
summary(dt2)
dim(dt2)

dt3<-read.table("/Users/mohsennabian/Desktop/MHEALTHDATASET/mHealth_subject3.log")
summary(dt3)
dim(dt3)

dt4<-read.table("/Users/mohsennabian/Desktop/MHEALTHDATASET/mHealth_subject4.log")
summary(dt4)
dim(dt4)

dt5<-read.table("/Users/mohsennabian/Desktop/MHEALTHDATASET/mHealth_subject5.log")
summary(dt5)
dim(dt5)

dt6<-read.table("/Users/mohsennabian/Desktop/MHEALTHDATASET/mHealth_subject6.log")
summary(dt6)
dim(dt6)

dt7<-read.table("/Users/mohsennabian/Desktop/MHEALTHDATASET/mHealth_subject7.log")
summary(dt7)
dim(dt7)

dt8<-read.table("/Users/mohsennabian/Desktop/MHEALTHDATASET/mHealth_subject8.log")
summary(dt8)
dim(dt8)

dt9<-read.table("/Users/mohsennabian/Desktop/MHEALTHDATASET/mHealth_subject9.log")
summary(dt9)
dim(dt9)

dt10<-read.table("/Users/mohsennabian/Desktop/MHEALTHDATASET/mHealth_subject10.log")
summary(dt10)
dim(dt10)

all<-rbind(dt1,dt2,dt3,dt4,dt5,dt6,dt7,dt8,dt9,dt10)


names(all)[24]<-"Y"
names(all)

all$Y<-as.factor(all$Y)
summary(all)
str(all)

# So 'all' is the all the data we have. 23 attributes (column 1-23) and 1 label(column 24) with 13 classes. 

```


2) Normalizing data:

```{r,results="hide"}

library('som')
all_norm<-normalize(all[,-24], byrow=FALSE)
all_norm<-as.data.frame(all_norm)
all_norm$Y<-all$Y
head(all_norm)

#'all_norm' is the 'all' data frame but normalized. 


dt<-all_norm[sample(nrow(all_norm), nrow(all_norm)), ]

#dt is the shuffled format of 'all_norm' 

```



3) Specifying Training and Test data


```{r,eval=FALSE}

ntr<-30000 #training
nte<-1000 #test

r<-0.8  #percentage of training data to be selected from rare activities. 

```


```{r,eval=FALSE}
train1<-dt[1:ntr,]
dim(train1)

test1<-dt[(ntr+1):(ntr+nte),]
dim(test1)
```

4) Applying Bagging method

###Bagging 
```{r,eval=FALSE}
# load the package
library(ipred)

fit1 <- bagging(Y~., data=train1)
ypred_bg1<- predict(fit1, test1[,-24], type="class")


#Evaluation:

true_rare(test1,ypred_bg1)
true_zero(test1,ypred_bg1)

```

Bagging does a very good job on classifying the activities based on the sensor data. In fact after Random Forrest ,SVM and Boosting, Baging was the most successful classifier in my dataset. 
