---
title: "HW12_2"
output: html_document
---

Answer the following questions:   
Q1)  Can some form of Social Network analysis help in your research project?    
Social Network analysis can be a part of my research, however, the paper is focused on supervised classificattion learning. Thus, a comprehensive study on network analysis is not possible. 

therefore, we used Michael Hahsler's network data to make somehow a comperhensove study to familiarize ourselfs with different aspects of social network analysis in R.

```{r}
set.seed(1234)

library(igraph)
source("http://michael.hahsler.net/SMU/ScientificCompR/code/map.R")
```

#Create a random graph
```{r}
g <-erdos.renyi.game(50, 3/50)
summary(g)
```
#Plot graphs
```{r}
plot(g)
```

Plot with an externally created layout
```{r}
layout <-layout.fruchterman.reingold(g)
plot(g, layout=layout)
```


Interactive plot

```{r}
tkplot(g)
```

Inspect vertex degree and use the degree for vertex size
```{r}
degree(g)



hist(degree(g))

plot(g, layout=layout, vertex.size=map(degree(g),c(1,20)))

```


Number of paths between vertex 1 and 2
```{r}
edge.disjoint.paths(g, 1, 2)
```
Betweenness centrality - number of geodesics (shortest paths) going through a vertex or an edge

```{r}
betweenness(g)

edge.betweenness(g)

plot(g, layout=layout,
  vertex.size=map(betweenness(g),c(1,15)),
  edge.width=map(edge.betweenness(g), c(1,10)))
```

Diameter of graph - Length of the longest geodesic path

```{r}
diameter(g)

get.diameter(g)
```

Maximal independent vertex sets
```{r}
independence.number(g)

col <- rep("blue",length(V(g)))
col[maximal.independent.vertex.sets(g)[[1]]] <- "red"
plot(g, layout=layout, vertex.color=col)
```


Find connected components


```{r}

cl <- clusters(g)
cl

plot(g, layout=layout, vertex.color=cl$membership+1L)



```


Get connected components
```{r}
dg <- decompose.graph(g)
length(dg)

plot(dg[[1]])
```

plot of chunk unnamed-chunk-12
```{r}
plot(dg[[2]])
```


plot of chunk unnamed-chunk-12

Max-flow between vertices 1 and 2

```{r}
graph.maxflow(g, 1, 2)
```

Min cut (of the first connected component)
```{r}
graph.mincut(dg[[1]], value.only=FALSE)
col <- rep("blue", length(V(dg[[1]])))
col[graph.mincut(dg[[1]], value.only=FALSE)[["partition2"]]] <- "red"
plot(dg[[1]], vertex.color=col)
```

plot of chunk unnamed-chunk-14

Find neighborhood of vertices

```{r}
gn <- graph.neighborhood(g, order = 1)
plot(gn[[1]])

```


plot of chunk unnamed-chunk-15
```{r}
plot(gn[[2]])

```
plot of chunk unnamed-chunk-15

Find cliques

```{r}
mc <- maximal.cliques(g)
length(mc)

```

click size

```{r}
sapply(mc, length)

col <- rep("blue", length(V(g)))
```
plot the largest (last) click
```{r}
col[mc[[length(mc)]]] <- "red"

plot(g, layout=layout, vertex.color=col)
```
plot of chunk unnamed-chunk-18
Large graph (social networks)


```{r}
g <- barabasi.game(1000, power=1)
layout <- layout.fruchterman.reingold(g)
plot(g, layout=layout, vertex.size=2,
  vertex.label=NA, edge.arrow.size=.2)
```

plot of chunk unnamed-chunk-19

Page Rank

```{r}
pr <- page.rank(g)$vector
plot(g, layout=layout, vertex.size=map(pr, c(1,20)),
  vertex.label=NA, edge.arrow.size=.2)
plot of chunk unnamed-chunk-20
```

Eigenvector centrality


```{r}
ec <- evcent(g)$vector
plot(g, layout=layout, vertex.size=map(ec, c(1,20)), vertex.label=NA, edge.arrow.size=.2)
```
plot of chunk unnamed-chunk-21

Kleinberg’s hub and authority scores

```{r}
auth <- authority.score(g)$vector
hub <- hub.score(g)$vector
plot(g, layout=layout, vertex.size=map(hub, c(1,5)), vertex.label=NA, edge.arrow.size=.2)

```

plot of chunk unnamed-chunk-22

```{r}
plot(g, layout=layout, vertex.size=map(auth, c(1,20)), vertex.label=NA, edge.arrow.size=.2)

```

plot of chunk unnamed-chunk-22

Community detection (betweenness)

```{r}
eb <- edge.betweenness.community(g)
```

create 10 communities

```{r}
member <- community.to.membership(g, eb$merges,step=nrow(eb$merges)-10L+1L)
plot(g,vertex.color= rainbow(10, .8, .8, alpha=.8)[member$membership+1L],
  vertex.size=5, layout=layout,  vertex.label=NA,
  edge.arrow.size=.2)
```

plot of chunk unnamed-chunk-24

Random walk

```{r}
eb <- walktrap.community(g)
```

create 10 communities
```{r}
member <- community.to.membership(g, eb$merges,
  step=nrow(eb$merges)-10L+1L)
plot(g,
  vertex.color= rainbow(10, .8, .8, alpha=.8)[member$membership+1L],
  vertex.size=5, layout=layout,  vertex.label=NA,
  edge.arrow.size=.2)
```

