# Assingment

* Use a genetic algorithm to search for a solution to the famous Travelling Salesman Problem in R. The same problem that we used in the simulated annealing lesson. 

###################################
Q1)  How does the genetic algorithm solution compare to the simulated annealing solution? 

here is the genetic algorithm for traveller salesaman:


```{r}
library(GA)
data("eurodist", package = "datasets")


 D <- as.matrix(eurodist)
 
 

tourLength <- function(tour, distMatrix) 
{
    tour <- c(tour, tour[1])
    route <- embed(tour, 2)[,2:1]
    sum(distMatrix[route])
}

tspFitness <- function(tour, ...) 1/tourLength(tour, ...)

GA <- ga(type = "permutation", fitness = tspFitness, distMatrix = D,min = 1, max = attr(eurodist, "Size"), popSize = 50, maxiter = 5000,run = 500,pcrossover = 0.8, pmutation = 0.2)
 
 
summary(GA)
apply(GA@solution, 1, tourLength, D)
 
mds <- cmdscale(eurodist)
mds 
x <- mds[, 1]
y <- -mds[, 2]
plot(x, y, type = "n", asp = 1, xlab = "", ylab = "")
abline(h = pretty(range(x),10),v=pretty(range(y),10),col = "light gray")

tour <- GA@solution[1, ]

tour <- c(tour, tour[1])

n <- length(tour)

arrows(x[tour[-n]], y[tour[-n]], x[tour[-1]],y[tour[-1]],length = 0.15, angle = 25, col = "steelblue",lwd = 2)
text(x, y, labels(eurodist), cex=0.8)
   

```

As we see with the default pcrossover=0.8 and pmutation 0.2 we end up a rout with the distance=12893 . The path is starting with : 10-20-7-11-3-4-18 .... and ends up to: 19-1-21-17-6-10     



Lets run simulated anealing once again to see the result.




```{r}

# Distance function
distance <- function(sq) 
  {  # Target function
  sq2 <- embed(sq, 2)
  return(as.numeric(sum(CityDistMatx[cbind(sq2[,2],sq2[,1])])))
 }


# Generate new candidates
GenSeq <- function(sq)
{  # Generate new candidate sequence
  idx <- seq(2, NROW(CityDistMatx)-1, by=1)
  ChangePoints <- sample(idx, size=2, replace=FALSE)
  tmp <- sq[ChangePoints[1]]
  sq[ChangePoints[1]] <- sq[ChangePoints[2]]
  sq[ChangePoints[2]] <- tmp
  return(as.numeric(sq))
}



CityDistMatx <- as.matrix(eurodist)
CityDistMatx

cities<-labels(eurodist)
cities

initial.tour <- c(1,2:NROW(CityDistMatx),1)  
# Initial sequence
initial.tour
initial.d<-distance(initial.tour)
initial.d
for(i in 1:length(initial.tour))
{
  print(cities[initial.tour[i]])  
}  
set.seed(333) # chosen to get a good soln relatively quickly

#  box-constrained optimization and simulated annealing
# method = "SANN" performs simulated annealing
# Method "SANN" is by default a variant of simulated annealing given in Belisle (1992)
res <- optim(initial.tour, distance, GenSeq, method = "SANN",
control = list(maxit = 30000, temp = 2000,trace=TRUE,REPORT = 500))

res  # Near optimum distance around 12842
final.tour<-res$par
final.tour
final.d<-distance(final.tour)
final.d
initial.d
final.d/initial.d


cities.xy <- cmdscale(eurodist)
cities.xy
for(i in 1:length(final.tour))
{
  print(cities[final.tour[i]])  
}  
rx <- range(x <- cities.xy[,1])
ry <- range(y <- -cities.xy[,2])
rx
ry
x
y
initial.tour
## remove last element to draw arrows from point to point
s <-head(initial.tour, -1)
s
plot(x, y, type="n", asp=1, xlab="", ylab="", main="initial solution of traveling salesman problem")
arrows(x[s], y[s], x[s+1], y[s+1], col="green")
text(x, y, labels(cities), cex=0.8)
plot(x, y, type="n", asp=1, xlab="", ylab="", main="initial solution of traveling salesman problem")
arrows(x[s], y[s], x[s+1], y[s+1], col="green")
text(x, y, labels(cities), cex=0.8)
text(x, y, labels(eurodist), cex=0.8)
final.tour
## draw lines from point to point
s <-head(final.tour, -1)
s
df = data.frame(x[s],y[s])
df
plot(x, y, type="n", asp=1, xlab="", ylab="", main="optimized simulated annealing traveling salesman problem")
lines(df$x, df$y, col="red")
text(x, y, labels(cities), cex=0.8)
plot(x, y, type="n", asp=1, xlab="", ylab="", main="optimized simulated annealing traveling salesman problem")
lines(df$x, df$y, col="red")
text(x, y, labels(cities), cex=0.8)
text(x, y, labels(eurodist), cex=0.8)
```

Here after 30000 iteration we end up the path length=12907
As we see, we have slighty better result using genetic algorithm optimization. 
I believe since simulated anealing is converging, increasing the maxitereation might end up to better solution. The point here is that both techniques are fast and leads to a very close and great solution. 


############################################################

Q2)  How does the crossover probability effect the genetic algorithm solution?  

Now lets play with the crossover probability.
We write a function to return the distance of the resutatant path for the given crossover 
probability. 
and finally plot the distance with respect to the crossovers.




```{r}

final_dist<-function (crossover)
{
  data("eurodist", package = "datasets")


 D <- as.matrix(eurodist)
 


tspFitness <- function(tour, ...) 1/tourLength(tour, ...)

GA <- ga(type = "permutation", fitness = tspFitness, distMatrix = D,min = 1, max = attr(eurodist, "Size"), popSize = 50, maxiter = 5000,run = 500,pcrossover = crossover, pmutation = 0.2)
 
 
summary(GA)
return(apply(GA@solution, 1, tourLength, D)[1])


}
j=1
for (i in c(seq(0.1, 1, 0.02)))
{
  x[j]=i;
  y[j]=final_dist(i)
  j=j+1
  print (j)
}
plot(x,y)
```

The plot indicates us that mostly higher crossover around 0.7 to 1 is a decent value, however, lower crossover values has tendency to worsen the result.



####################################################

Q3)   How does the mutation probability effect the genetic algorithm solution?     


Here, the same approach is utilized by writing a distance function and a loop over the 
range of 
```{r}
final_dist2<-function (mutation)
{
  data("eurodist", package = "datasets")


 D <- as.matrix(eurodist)
 


tspFitness <- function(tour, ...) 1/tourLength(tour, ...)

GA <- ga(type = "permutation", fitness = tspFitness, distMatrix = D,min = 1, max = attr(eurodist, "Size"), popSize = 50, maxiter = 5000,run = 500,pcrossover = 0.8, pmutation = mutation)
 
 
summary(GA)
return(apply(GA@solution, 1, tourLength, D)[1])


}
j=1
for (i in c(seq(0.0, 1, 0.02)))
{
  x[j]=i;
  y[j]=final_dist2(i)
  j=j+1
  print (j)
}
plot(x,y)

```

The result indicates that mutation value is mostly providing better solution in the 
range of 0.1 to 0.4 . However, increasing mutation like 0.6 to 1, will likely decreases the 
performance of the algorithm.







