# Assingment

Use simulated annealing to search for a solution to the famous Travelling Salesman Problem in R.  




######################################################################

       





```{r}
#install.packages("ggplot2")
#install.packages("stats")
require(ggplot2)
require(stats)
```


```{r}

# Distance function
distance <- function(sq) 
  {  # Target function
  sq2 <- embed(sq, 2)
  return(as.numeric(sum(CityDistMatx[cbind(sq2[,2],sq2[,1])])))
 }


# Generate new candidates
GenSeq <- function(sq) {  # Generate new candidate sequence
  idx <- seq(2, NROW(CityDistMatx)-1, by=1)
  ChangePoints <- sample(idx, size=2, replace=FALSE)
  tmp <- sq[ChangePoints[1]]
  sq[ChangePoints[1]] <- sq[ChangePoints[2]]
  sq[ChangePoints[2]] <- tmp
  return(as.numeric(sq))
}

```

In the begining, we use uclidian distance for future use when are comparing 

```{r}

CityDistMatx <- as.matrix(eurodist)
CityDistMatx

cities<-labels(eurodist)
cities

initial.tour <- c(1,2:NROW(CityDistMatx),1)  
# Initial sequence
initial.tour
initial.d<-distance(initial.tour)
initial.d
for(i in 1:length(initial.tour))
{
  print(cities[initial.tour[i]])  
}  
set.seed(333) # chosen to get a good soln relatively quickly

#  box-constrained optimization and simulated annealing
# method = "SANN" performs simulated annealing
# Method "SANN" is by default a variant of simulated annealing given in Belisle (1992)
res <- optim(initial.tour, distance, GenSeq, method = "SANN",
control = list(maxit = 30000, temp = 2000,trace=TRUE,REPORT = 500))

res  # Near optimum distance around 12842
final.tour<-res$par
final.tour
final.d<-distance(final.tour)
final.d
initial.d
final.d/initial.d


cities.xy <- cmdscale(eurodist)
cities.xy
for(i in 1:length(final.tour))
{
  print(cities[final.tour[i]])  
}  
rx <- range(x <- cities.xy[,1])
ry <- range(y <- -cities.xy[,2])
rx
ry
x
y
initial.tour
## remove last element to draw arrows from point to point
s <-head(initial.tour, -1)
s
plot(x, y, type="n", asp=1, xlab="", ylab="", main="initial solution of traveling salesman problem")
arrows(x[s], y[s], x[s+1], y[s+1], col="green")
text(x, y, labels(cities), cex=0.8)
plot(x, y, type="n", asp=1, xlab="", ylab="", main="initial solution of traveling salesman problem")
arrows(x[s], y[s], x[s+1], y[s+1], col="green")
text(x, y, labels(cities), cex=0.8)
text(x, y, labels(eurodist), cex=0.8)
final.tour
## draw lines from point to point
s <-head(final.tour, -1)
s
df = data.frame(x[s],y[s])
df
plot(x, y, type="n", asp=1, xlab="", ylab="", main="optimized simulated annealing traveling salesman problem")
lines(df$x, df$y, col="red")
text(x, y, labels(cities), cex=0.8)
plot(x, y, type="n", asp=1, xlab="", ylab="", main="optimized simulated annealing traveling salesman problem")
lines(df$x, df$y, col="red")
text(x, y, labels(cities), cex=0.8)
text(x, y, labels(eurodist), cex=0.8)
```



 Answer the following questions:  
 Q1)  Write a different distance function. How does the effect the final solution?  



Here is a Manhatan Distance I wrote:

```{r}

# Creating Manhatan matrix

cities.xy <- cmdscale(eurodist)
cities.xy<-(cities.xy)

ManhDistmat<-matrix(, nrow = nrow(cities.xy), ncol =nrow(cities.xy))

for (i in 1:nrow(cities.xy))
{ 
  for (j in 1:nrow(cities.xy))
  {
    ManhDistmat[i,j]=as.numeric(abs(cities.xy[i,1]-cities.xy[j,1]))+as.numeric(abs(cities.xy[i,2]-cities.xy[j,2]))
    
  }
}


# Manhatan Distance function
distance2 <- function(sq) 
  {  # Target function
  sq2 <- embed(sq, 2)
  return(as.numeric(sum(ManhDistmat[cbind(sq2[,2],sq2[,1])])))
 }


```


Now we use manhatan distance function to see how the results are different from 
Eculucidian distance.



```{r}

CityDistMatx <- as.matrix(eurodist)
CityDistMatx

cities<-labels(eurodist)
cities

initial.tour <- c(1,2:NROW(CityDistMatx),1)  
# Initial sequence
initial.tour
initial.d<-distance2(initial.tour)
initial.d
for(i in 1:length(initial.tour))
{
  print(cities[initial.tour[i]])  
}  
set.seed(333) # chosen to get a good soln relatively quickly

#  box-constrained optimization and simulated annealing
# method = "SANN" performs simulated annealing
# Method "SANN" is by default a variant of simulated annealing given in Belisle (1992)
res <- optim(initial.tour, distance2, GenSeq, method = "SANN",
control = list(maxit = 30000, temp = 2000,trace=TRUE,REPORT = 500))

res  # Near optimum distance around 12842
final.tour<-res$par
final.tour
final.d<-distance2(final.tour)
final.d
initial.d
final.d/initial.d


cities.xy <- cmdscale(eurodist)
cities.xy
for(i in 1:length(final.tour))
{
  print(cities[final.tour[i]])  
}  
rx <- range(x <- cities.xy[,1])
ry <- range(y <- -cities.xy[,2])
rx
ry
x
y
initial.tour
## remove last element to draw arrows from point to point
s <-head(initial.tour, -1)
s
plot(x, y, type="n", asp=1, xlab="", ylab="", main="initial solution of traveling salesman problem")
arrows(x[s], y[s], x[s+1], y[s+1], col="green")
text(x, y, labels(cities), cex=0.8)
plot(x, y, type="n", asp=1, xlab="", ylab="", main="initial solution of traveling salesman problem")
arrows(x[s], y[s], x[s+1], y[s+1], col="green")
text(x, y, labels(cities), cex=0.8)
text(x, y, labels(eurodist), cex=0.8)
final.tour
## draw lines from point to point
s <-head(final.tour, -1)
s
df = data.frame(x[s],y[s])
df
plot(x, y, type="n", asp=1, xlab="", ylab="", main="optimized simulated annealing traveling salesman problem")
lines(df$x, df$y, col="red")
text(x, y, labels(cities), cex=0.8)
plot(x, y, type="n", asp=1, xlab="", ylab="", main="optimized simulated annealing traveling salesman problem")
lines(df$x, df$y, col="red")
text(x, y, labels(cities), cex=0.8)
text(x, y, labels(eurodist), cex=0.8)
```


As we can see using Manhatan Distance function results in  18387 distance compared to 12907 which resulted from Eulicidian Distance. 






#########################################################################

Q2)  * Does scaling, normalization or leaving the data unscaled make a difference for simulated annealing? 



```{r}

CityDistMatx <- as.matrix(eurodist)
CityDistMatx<-scale(CityDistMatx)

cities<-labels(eurodist)
cities

initial.tour <- c(1,2:NROW(CityDistMatx),1)  
# Initial sequence
initial.tour
initial.d<-distance(initial.tour)
initial.d
for(i in 1:length(initial.tour))
{
  print(cities[initial.tour[i]])  
}  
set.seed(333) # chosen to get a good soln relatively quickly

#  box-constrained optimization and simulated annealing
# method = "SANN" performs simulated annealing
# Method "SANN" is by default a variant of simulated annealing given in Belisle (1992)
res <- optim(initial.tour, distance, GenSeq, method = "SANN",
control = list(maxit = 30000, temp = 2000,trace=TRUE,REPORT = 500))

res  # Near optimum distance around 12842
final.tour<-res$par
final.tour
final.d<-distance(final.tour)
final.d
initial.d
final.d/initial.d


cities.xy <- cmdscale(eurodist)
cities.xy
for(i in 1:length(final.tour))
{
  print(cities[final.tour[i]])  
}  
rx <- range(x <- cities.xy[,1])
ry <- range(y <- -cities.xy[,2])
rx
ry
x
y
initial.tour
## remove last element to draw arrows from point to point
s <-head(initial.tour, -1)
s
plot(x, y, type="n", asp=1, xlab="", ylab="", main="initial solution of traveling salesman problem")
arrows(x[s], y[s], x[s+1], y[s+1], col="green")
text(x, y, labels(cities), cex=0.8)
plot(x, y, type="n", asp=1, xlab="", ylab="", main="initial solution of traveling salesman problem")
arrows(x[s], y[s], x[s+1], y[s+1], col="green")
text(x, y, labels(cities), cex=0.8)
text(x, y, labels(eurodist), cex=0.8)
final.tour
## draw lines from point to point
s <-head(final.tour, -1)
s
df = data.frame(x[s],y[s])
df
plot(x, y, type="n", asp=1, xlab="", ylab="", main="optimized simulated annealing traveling salesman problem")
lines(df$x, df$y, col="red")
text(x, y, labels(cities), cex=0.8)
plot(x, y, type="n", asp=1, xlab="", ylab="", main="optimized simulated annealing traveling salesman problem")
lines(df$x, df$y, col="red")
text(x, y, labels(cities), cex=0.8)
text(x, y, labels(eurodist), cex=0.8)

```

The path we end up after scaling is different from unscaled distances. In fact the path is screwed up and non-sense which leads us to conclude that for path optimization, scaling is NOT a good idea. 





############################################################


Q3) Try a few differnt initial tours. How does the effect the final solution?


```{r}


CityDistMatx <- as.matrix(eurodist)
CityDistMatx

cities<-labels(eurodist)
cities

# Randomizing the initial path: 

initial.tour <- c(1,sample(2:NROW(CityDistMatx), NROW(CityDistMatx)-1),1)  
# Initial sequence
initial.tour
initial.d<-distance(initial.tour)
initial.d
for(i in 1:length(initial.tour))
{
  print(cities[initial.tour[i]])  
}  
set.seed(333) # chosen to get a good soln relatively quickly

#  box-constrained optimization and simulated annealing
# method = "SANN" performs simulated annealing
# Method "SANN" is by default a variant of simulated annealing given in Belisle (1992)
res <- optim(initial.tour, distance, GenSeq, method = "SANN",
control = list(maxit = 30000, temp = 2000,trace=TRUE,REPORT = 500))

res  # Near optimum distance around 12842
final.tour<-res$par
final.tour
final.d<-distance(final.tour)
final.d
initial.d
final.d/initial.d


cities.xy <- cmdscale(eurodist)
cities.xy
for(i in 1:length(final.tour))
{
  print(cities[final.tour[i]])  
}  
rx <- range(x <- cities.xy[,1])
ry <- range(y <- -cities.xy[,2])
rx
ry
x
y
initial.tour
## remove last element to draw arrows from point to point
s <-head(initial.tour, -1)
s
plot(x, y, type="n", asp=1, xlab="", ylab="", main="initial solution of traveling salesman problem")
arrows(x[s], y[s], x[s+1], y[s+1], col="green")
text(x, y, labels(cities), cex=0.8)
plot(x, y, type="n", asp=1, xlab="", ylab="", main="initial solution of traveling salesman problem")
arrows(x[s], y[s], x[s+1], y[s+1], col="green")
text(x, y, labels(cities), cex=0.8)
text(x, y, labels(eurodist), cex=0.8)
final.tour
## draw lines from point to point
s <-head(final.tour, -1)
s
df = data.frame(x[s],y[s])
df
plot(x, y, type="n", asp=1, xlab="", ylab="", main="optimized simulated annealing traveling salesman problem")
lines(df$x, df$y, col="red")
text(x, y, labels(cities), cex=0.8)
plot(x, y, type="n", asp=1, xlab="", ylab="", main="optimized simulated annealing traveling salesman problem")
lines(df$x, df$y, col="red")
text(x, y, labels(cities), cex=0.8)
text(x, y, labels(eurodist), cex=0.8)


```


As we see by setting a random initial route we will end up a different result
. The total distance with this random initialization is 14862 as opposed to the previous manual initialization  which led the path length 12907.
However, If we let the algorithm converges, we will end up all the same value
of total distance and same route. 








