---
title: "Homework 1_a"
author: "Mohsen Nabian"
date: "9/28/2015"
output: html_document
---

## Homework Question 1

_Probability distributions_

* Replicate and plot the fat-tailed Cauchy distributions from https://en.wikipedia.org/wiki/Cauchy_distribution
* Load the file M01_Lesson_02_Q1.csv
* Answer the following questions for the data in each column:
    * How is the data distributed?
    * What are the summary statistics?
    * Are there anomalies/outliers?
    * Try to regenerate the data in each column.
    * Plot your regenerated data versus the original data using a faceted graph. How does it compare? 



```{r}
# Drawing the Cauchy Distribution
#install.packages('ggplot2')
require(ggplot2)
##############
```




```{r}
cauchy<-data.frame(A=rcauchy(100000,0,0.5),B=rcauchy(100000,0,1),
                   C=rcauchy(100000,0,2),D=rcauchy(100000,-2,1))

require(reshape2)
rnd <- melt(data=cauchy)    #Changing the form of data frame
head(rnd)
g<-ggplot(data=rnd)+ aes(x=value)+aes(group=variable)+aes(color=variable)
g<-g+geom_density()
g<-g+xlim(-5,5)+ylim(0,0.7)
g<-g+labs(title="Wikipedia Cauchy Distributions", y="P(x)", x="x")
g


#####################################################
```






```{r}

DATA<-read.csv("C:/Users/nabian.m/OneDrive/Fall 2015/intr. Data mining and Machine learning/Week1/HomeWork/M01_Lesson_02_Q1.csv",header=TRUE)
head(DATA)
```


##Column A

```{r}
#Column A

g0<-ggplot(data=DATA)
g1<-g0+aes(x=A)+geom_density()
g1
vect<-DATA$A
summary(vect)
```


based on the shape of the distribution and the fact that mean and median are very close, the distribution is very  similar to a gussian distribution.


```{r}
lowerq = quantile(vect)[2]
upperq = quantile(vect)[4]
innerq = upperq - lowerq #Or use IQR(data)
threshold.upper = (innerq * 1.5) + upperq
threshold.lower = lowerq - (innerq * 1.5)
outlier<-sum((vect>threshold.upper))+sum((vect<threshold.lower))
outlier
##############################
```

So there is only 1 outlier here based on this definition that we coded.

```{r}

miu<-mean(DATA$A)
sdd<-sd(DATA$A)
my_A<-rnorm(length(DATA$A),miu,sdd)
df<-data.frame(x=DATA$A,Type=as.factor('Real'))
df<-rbind(df,data.frame(x=my_A,Type=as.factor('Predicted')))

g0<-ggplot(data=df)+aes(x=x,color=Type)+geom_density()
g0

```


##Column B
```{r}
#Column B

g0<-ggplot(data=DATA)
g2<-g0+aes(x=B)+geom_density()
g2
vect<-DATA$B
summary(vect)
```

This is also close to the normal distribution due to the shape as well as the fact that median and mean are very close. However, a handful of outliers are visually detectable from the distribution.


```{r}
lowerq = quantile(vect)[2]
upperq = quantile(vect)[4]
innerq = upperq - lowerq #Or use IQR(data)
threshold.upper = (innerq * 1.5) + upperq
threshold.lower = lowerq - (innerq * 1.5)
outlier<-sum((vect>threshold.upper))+sum((vect<threshold.lower))
outlier
##############################

```
So there are 3 outliers here based on this definition that we coded.

```{r}

miu<-mean(DATA$B)
sdd<-sd(DATA$B)
my_B<-rnorm(length(DATA$B),miu,sdd)
df<-data.frame(x=DATA$B,Type=as.factor('Real'))
df<-rbind(df,data.frame(x=my_B,Type=as.factor('Predicted')))

g0<-ggplot(data=df)+aes(x=x,color=Type)+geom_density()
g0


```




##Column C
```{r}
#Column C

g0<-ggplot(data=DATA)
g3<-g0+aes(x=C)+geom_density()
g3
vect<-DATA$C
summary(vect)
```

Based on the fact that the mean is slightly larger than the median, we can say this distribution is slightly right skewed.
However, we can not visually detect that. Most of the values are located between 5 and 8. 

```{r}
lowerq = quantile(vect)[2]
upperq = quantile(vect)[4]
innerq = upperq - lowerq #Or use IQR(data)
threshold.upper = (innerq * 1.5) + upperq
threshold.lower = lowerq - (innerq * 1.5)
outlier<-sum((vect>threshold.upper))+sum((vect<threshold.lower))
outlier
###############################
```
Based on the outlier definition that is coded, there is no outlier.

```{r}


miu<-mean(DATA$C)
sdd<-sd(DATA$C)
my_C<-rnorm(length(DATA$C),miu,sdd)
df<-data.frame(x=DATA$C,Type=as.factor('Real'))
df<-rbind(df,data.frame(x=my_C,Type=as.factor('Predicted')))

g0<-ggplot(data=df)+aes(x=x,color=Type)+geom_density()
g0

```


##Column D
```{r}
#Column D

g0<-ggplot(data=DATA)
g4<-g0+aes(x=D)+geom_density()
g4
vect<-DATA$D
summary(vect)
```

The mean and median are very close an d the distribution look like a normal distribution. However, we might have a few outliers.


```{r}
lowerq = quantile(vect)[2]
upperq = quantile(vect)[4]
innerq = upperq - lowerq #Or use IQR(data)
threshold.upper = (innerq * 1.5) + upperq
threshold.lower = lowerq - (innerq * 1.5)
outlier<-sum((vect>threshold.upper))+sum((vect<threshold.lower))
outlier
###############################
```
we have 2 outliers based on the definition.

```{r}
miu<-mean(DATA$D)
sdd<-sd(DATA$D)
my_D<-rnorm(length(DATA$D),miu,sdd)
df<-data.frame(x=DATA$D,Type=as.factor('Real'))
df<-rbind(df,data.frame(x=my_D,Type=as.factor('Predicted')))

g0<-ggplot(data=df)+aes(x=x,color=Type)+geom_density()
g0

```


##Column E
```{r}
#Column E

g0<-ggplot(data=DATA)
g5<-g0+aes(x=E)+geom_density()
g5
vect<-DATA$E
summary(vect)
```
Here the distribution is close to a uniform distribution. 


```{r}
lowerq = quantile(vect)[2]
upperq = quantile(vect)[4]
innerq = upperq - lowerq #Or use IQR(data)
threshold.upper = (innerq * 1.5) + upperq
threshold.lower = lowerq - (innerq * 1.5)
outlier<-sum((vect>threshold.upper))+sum((vect<threshold.lower))
outlier
###############################

```
No outlier found. 


```{r}



miu<-mean(DATA$E)

my_E<-runif(length(DATA$E),50,330)
df<-data.frame(x=DATA$E,Type=as.factor('Real'))
df<-rbind(df,data.frame(x=my_E,Type=as.factor('Predicted')))

g0<-ggplot(data=df)+aes(x=x,color=Type)+geom_density()
g0



```

