---
title: "Homework 1_b"
author: "Mohsen Nabian"
date: "9/28/2015"
output: html_document
---



##First, load the file M01_quasi_twitter.csv 
Next, answer the following questions for the data in each column:

###1)How is the data distributed?
###2)What happens when you test distribution assumptions (e.g. normal distributions or skewed, etc)?
###3)What are the summary statistics?
###4)Are there anomalies/outliers?
###5)Can you identify the following:
  useful raw data and transforms (e.g. log(x))
###6)data quality problems
###7)outliers
###8)subsets of interest
###9)Finally, suggest any functional relationships.



##Solution: 

first to install the required packages
```{r}

require(ggplot2)
```

Now we write the outlier function as the following. 
This function will find the outlier of any given vector.

```{r}
Outlier = function(vector)
{
  lowerq = quantile(vector)[2]
  upperq = quantile(vector)[4]
  innerq = upperq - lowerq #Or use IQR(data)
  threshold.upper = (innerq * 1.5) + upperq
  threshold.lower = lowerq - (innerq * 1.5)
  NO_outlier<-sum((vector>threshold.upper))+sum((vector<threshold.lower))
  print(NO_outlier)
  outliers<-vector[(vector>threshold.upper)|(vector<threshold.lower)]
  return(outliers)
}
```


Now we read the data into the twitter data frame: 
```{r}
twitter<-read.csv(file='C:/Users/nabian.m/Desktop/M01_quasi_twitter.csv',header=TRUE)
twitter<-na.omit(twitter)
dim(twitter)
head(twitter)
```


#1st Column ="Screen Name"
```{r}
class(twitter$screen_name)
screen<-as.data.frame(table(twitter$screen_name))
dim(screen)
head(screen)
summary(screen$Freq)
```

So this means we have all frequencies=1. Which means for each screen name we have only one tweet. Which in fact is what we need to do an unbiased study.
Since we have 21916 factors with frequencies=1 it does no make sence to plot them. 

#2nd Column= created_at_month

```{r}
class(twitter$created_at_month)
vector<-twitter$created_at_month
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()
g

summary(vector)
```

since median is slightly smaller than the mean, that means the data a slightly right skewed. 
My expectation was a uniform distribution over the months, however, we see in months 3 and 4 we recieve about 50% to 70 % more frequent tweet. 
from the month 5 to 12 the distribution is almost uniform.

```{r}

Outlier(vector)
#as we expected there is no outlier.

```

#3rd Column= created_at_day
```{r}
class(twitter$created_at_day)
vector<-twitter$created_at_day
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()
g
summary(vector)
Outlier(vector)
```
So as we see it is almost a uniform distribution.
This distribution is more like a uniform distribution
#4th Column= created_at_year
```{r}
class(twitter$created_at_year)
vector<-twitter$created_at_year
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()+xlim(2006,2015)
g
```

the data collected from twitter is not uniform in terms of years. In fact it has the most amount of data from 2009 and least amount in 2007. From 2010 to 2015 data is almost uniform with the exception of 2015 which is slightly higher.

```{r}
summary(vector)
Outlier(vector)
 
```
meadian and mean are the same. So there is no skewness.
There is no outlier for year as expected.

#5th Column= country
```{r}
country<-as.data.frame(table(twitter$country))
country <- country[order(country$Freq),]
head(country)
tail(country)
dim(country)
summary(country$Freq)
```



as we see the results are intensively biased toward USA having 14905 tweets in comparision to the 2nd most frequent conuntry (Canada) with only 943 tweets.
 which is about  6% of U.S.A. All the tweets collected in the file is about 22000 and we have about 15000 from USA and 7000 from other countries mostly western developed countries.
So with a good approximation we can say the data belongs mostly to the US People.  


```{r}

g<-ggplot(data.frame(twitter$country))
g<-g+aes(x=twitter$country)+geom_histogram()
g
```
This plot also demonstrate this bais toward US users.




#7th Column= friends

```{r}
class(twitter$friends_count)
vector<-twitter$friends_count
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()+xlim(0,10000)
g
```


The plot demosntrate the a very high amount of users have friends less than 500. Which in fact make sense. 
We can also observe a acute increase in users having around 2000 connections. After that we see a smooth uniform and low value distribution. 
It is quite expected to have such left skewed plot for connections since most people are ordinary and not either so famous or social. 

This function equation is like y=a*exp(-bx)+c

```{r}
summary(vector)
```
Since it is the median is highly lower than the mean, the distribution is right skewed.

```{r}
Outlier(vector)
```

Based on the conventional definition of outlier, we have 2409 outliers in the file. However,since the distribution is highly left skewed.
we might change our definition. Moreover, the values in here is showing the reality of a network systems that a short percentage are dominantly attached to 
other nodes. As a result, It all depends on how we want to analyze and present the data. 




#8th Column= followers
```{r}
class(twitter$followers_count)
vector<-twitter$followers_count
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()+xlim(0,10000)
g
```

The plot demosntrate the a very high amount of users have followers less than 500. Which in fact make sense. 
# Based on our previous observation of friendship counts, I was completely expecting the same trend of having a sharply left skewed distribution.


This function equation is like y=a*exp(-bx)+c


```{r}
summary(vector)
```
Since it is the median is highly lower than the mean, the distribution is left skewed.
Outlier(vector)
Based on the conventional definition of outlier, we have 3026 outliers in the file. However,  since the distribution is highly left skewed 
 we might change our definition.So It all depends on how we want to analyze and present the data. 



#9th Column= statuses
```{r}
class(twitter$statuses_count)
vector<-twitter$statuses_count
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()+xlim(0,40000)
g
```

 This plot indicates that most of people have low statuses however, a few have so many. This is expected to be like this. 


```{r} 
summary(vector)
```
Since it is the median is highly lower than the mean, the distribution is right skewed.
```{r}
h<-Outlier(vector)
```
Based on the conventional definition of outlier, we have 2743 outliers in the file. However,  since the distribution is highly left skewed 
 we might change our definition.So It all depends on how we want to analyze and present the data. 


#10th Column = Favorites Count
```{r}
class(twitter$favourites_count)
vector<-twitter$favourites_count
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()+xlim(0,4000)
g

```
 This plot indicates that most of people have low Favorits count however, a few have so little. This is expected to be like this. 

```{r} 
summary(vector)
```
Since it is the median is highly lower than the mean, the distribution is right skewed.
```{r} 
h<-Outlier(vector)
```
Based on the conventional definition of outlier, we have 3031 outliers in the file. However,  since the distribution is highly left skewed 
 we might change our definition.So It all depends on how we want to analyze and present the data. 


#11th Column = favourited_count

```{r}
class(twitter$favourited_count)
vector<-twitter$favourited_count
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()+xlim(0,4000)
g
```

 This plot indicates that most of people have low Favorits count however, a few have so little. This is expected to be like this. 

```{r} 
summary(vector)
```
Since it is the median is highly lower than the mean, the distribution is right skewed.
```{r}
h<-Outlier(vector)
```
Based on the conventional definition of outlier, we have 3031 outliers in the file. However,  since the distribution is highly left skewed 
 we might change our definition.So It all depends on how we want to analyze and present the data. 

We may face data quality problem since many of our data points are 0 and quite a few are non zero. This will make the statistical analysis very hard. 


#12th Column = dob_day

```{r}
class(twitter$dob_day)
vector<-twitter$dob_day
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()+xlim(0,30)
g

```
 This plot indicates that most of people have low Favorits count however, a few have so little. This is expected to be like this. 

```{r} 
summary(vector)
```

Since it is the median is slightly  lower than the mean, the distribution is slightly right skewed.

```{r}
h<-Outlier(vector)
```

There is no outlier.

#############################################################

#13th Column = dob_year
```{r}
class(twitter$dob_year)
vector<-twitter$dob_year
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()+xlim(1900,2015)
g
```

 This plot indicates that most of people have low Favorits count however, a few have so little. This is expected to be like this. 

This function equation is like y=a*log(x)+c
```{r} 
summary(vector)
```
Since the median is higher than the mean, the distribution is left skewed.

```{r}
h<-Outlier(vector)
```
Based on the conventional definition of outlier, we have 3031 outliers in the file. However,  since the distribution is highly left skewed 
 we might change our definition.So It all depends on how we want to analyze and present the data. 

#########################################################

#14th Column = dob_month
```{r}
class(twitter$dob_month)
vector<-twitter$dob_month
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()+xlim(1,12)
g
```

 This plot indicates that most of people have low Favorits count however, a few have so little. This is expected to be like this. 
```{r}
summary(vector)
```
Since it is the median is highly lower than the mean, the distribution is right skewed.

```{r}
h<-Outlier(vector)

```
There are only 4 outliers. 



#15th Column = gender
```{r}
class(twitter$gender)
vector<-twitter$gender
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_histogram(coulor="blue")
g
```

 This plot indicates that most of people have low Favorits count however, a few have so little. This is expected to be like this. 
```{r}
summary(vector)
```
Since it is the median is highly lower than the mean, the distribution is left skewed.


#16th Column = mobile_favourites_count
```{r}
class(twitter$mobile_favourites_count)
vector<-twitter$mobile_favourites_count
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()
g
```

This plot indicates that most of people have 0 mobile Favorits count however, very few have huge number. This is expected to be like this. 
```{r}
summary(vector)
```
Since it is the median=0 and mean is much higher, the distribution is right skewed.

We may face data quality problem since many of our data points are 0 and quite a few are non zero. This will make the statistical analysis very hard. 



#17th Column = mobile_favourited_count
```{r}
class(twitter$mobile_favourited_count)
vector<-twitter$mobile_favourites_count
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()
g
```

This plot indicates that most of people have 0  Favorited count however, very few have huge number. This is expected to be like this.
```{r}
summary(vector)
```
Since it is the median=0 and mean is much higher, the distribution is right skewed.

```{r}
h<-Outlier(vector)

```
There are only many outliers based on the 1.5 coefficient. 
In fact they are not outliers, that is the nature of that variable to be highly skewed. 

We may face data quality problem since many of our data points are 0 and quite a few are non zero. This will make the statistical analysis very hard. 



#18th Column = education
```{r}
class(twitter$education)
vector<-twitter$education
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_histogram()+xlim(1,20)
g

```
This plot indicates that education is a aproximately a normal variable among people. 
```{r}
summary(vector)

```
mean is slightly smaller -> slightly left skewed. 

The distribution is very similar to normal distribution.

#19th Column = experience
```{r}
class(twitter$experience)
vector<-twitter$experience
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_histogram()+xlim(0,20)
g
```

This plot indicates that most of people have 0 or very low experience.
```{r}
summary(vector)
```
Since it is the median is highly lower than the mean, the distribution is right skewed.




#20th Column = age


```{r}
class(twitter$age)
vector<-twitter$age
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()+xlim(0,100)
g
```

As it could be seen, the age is almost normally distributed. 
```{r}
summary(vector)

```
mean and median are very close, NO skewness. 


#21th Column = race
```{r}
class(twitter$race)
vector<-twitter$race
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_histogram()
g
```

White race is h
```{r}
summary(vector)

```
Since it is the median is highly lower than the mean, the distribution is left skewed.

Based on the conventional definition of outlier, we have 3031 outliers in the file. However,  since the distribution is highly left skewed 
 we might change our definition.So It all depends on how we want to analyze and present the data. 

#22th column =wage
```{r}
class(twitter$wage)
vector<-twitter$wage
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()
g
```

 This plot indicates that most of people have low wages as opposed to minorities with high wages. We were expecting such distribution. 
```{r}
summary(vector)

```
Since it is the median is  lower than the mean, the distribution is right skewed.


```{r}
h<-Outlier(vector)

```
There are  many outliers based on the 1.5 coefficient. 
In fact they are not outliers, that is the nature wages  that  is to be highly skewed specially in United States. 



#23th Column =retweeted_count

```{r}
class(twitter$retweeted_count)
vector<-twitter$retweeted_count
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()
g
```

 This plot indicates that most of people have 0 retweeted counts however, very few have a little. This is expected to be like this. 
```{r}
summary(vector)

```
Since it is the median is highly lower than the mean, the distribution is left skewed.

```{r}
h<-Outlier(vector)

```
There are  many outliers based on the 1.5 coefficient. SInce we have many 0 values, cause to show that much outliers. 

We may face data quality problem since many of our data points are 0 and quite a few are non zero. This will make the statistical analysis very hard. 



#24th Column =retweet_count
```{r}
class(twitter$retweet_count)
vector<-twitter$retweet_count
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()
g
```

 This plot indicates that most of people have low Favorits count however, a few have so little. This is expected to be like this. 
```{r}
summary(vector)

```

Since it is the median is highly lower than the mean, the distribution is left skewed.
We may face data quality problem since many of our data points are 0 and quite a few are non zero. This will make the statistical analysis very hard. 

#25th Column =height



```{r}
class(twitter$height)
vector<-twitter$height
g<-ggplot(data.frame(vector))
g<-g+aes(x=vector)+geom_density()
g
```

 This plot indicates that height is almost a normal distribution.
```{r}
summary(vector)
```

mean and median are very close due to the normality of people's height. 

```{r}
h<-Outlier(vector)

```


---------

##Sub data 

We can extract different useful subsets: 
Like male and female:
```{r}
male<-subset(twitter,gender=='male')
head(male)
female<-subset(twitter,gender=='female')
head(female)
```
or we can extract USA users by:
```{r}
USA_users<-subset(twitter,country=='USA')
head(USA_users)
```


## Sample Regression

I am particularly interested in predicting of how many times a post is favourited.
SO I suggest the following linear model. 
```{r}
lm1<-lm(favourited_count~followers_count+statuses_count+friends_count,data=twitter)
summary(lm1)

```

As we see it turned out a linear model with all significant coefficient. In fact this model would make sense since these are all parameters that closely related and one affect the other. 

